## Demonstrating Model Validation on Credit Dataset

Libraries  
```{r}
library(tidyverse)
library(caret) #for splitting for validation
```

Read-in dataset  
```{r}
credit = read_csv("CreditData.csv")
```

Get rid of missing data rows 
```{r}
credit = credit %>% drop_na() #delete any row with an NA value
str(credit) #check structure after the drop
```

We will do pure k-fold cross-validation with no training and testing set. Can easily combine with train/test split. If you do this, do k-fold on training set.

Model with best single variable (by correlation).  
```{r}
ctrl = trainControl(method = "cv",number = 10) #set up caret 10 fold cross validation

set.seed(123) #set random number seed for cross validation
modCV = train(AnnualCharges ~ AnnualIncome, credit, method = "lm", trControl = ctrl, metric="Rsquared")
summary(modCV)
```

That's it. We are done. If we had done a train/test split, we would then evaluate the quality of the model on the testing set. 


